Formal Analysis of Rate-Monotonic Scheduling Implementation via Real-Time Maude

### 2015-10-08 RMS/submission.tex
finish completeness

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumerate}
\usepackage{epic}
\usepackage{pstricks}
%\usepackage{llncsdoc}

@@

\subsection{Specification}
A Real-Time Maude module specifies a \emph{real-time rewrite theory}
$(\Sigma, E\cup A , IR, TR)$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of possibly conditional equations, and $A$ a
  set of equational axioms such as associativiy, commutativity and
  identity.  $(\Sigma, E\cup A)$ specifies the system's state space as
  an algebraic data type, and includes a built-in specification of a
  sort \verb|Time|.
\item $IR$ is a set of \emph{labeled conditional rewrite rules}
  specifying the system's local transitions, each of which has the
  form $[l]~:~t\rightarrow t'\mbox{ \textbf{if}
  }\bigwedge^n_{j=1}cond_j$, where each $cond_j$ is an equality
  $u_j=v_j$, and $l$ is a \emph{label}. Such a rule specifies an
  \emph{instaneous transition} from an instance of $t$ to the
  corresponding instance of $t'$, \emph{provided} the conditions hold.
\item $TR$ is a set of \emph{tick rules} $[l]~:~\{t\}\rightarrow\{t'\}
  \mbox{ \textbf{in time} }\tau\mbox{ \textbf{if} }cond$ that advance
  time in the \emph{entire} state $t$ by $\tau$ time units.
\end{itemize}

@@

\subsection{Scenarios}
We use the following setting for our verification, which is from the 
statistics provided by our industrial partner:
\begin{itemize}
\item The interrupt cycle is $5ms$.
\item The scheduling time ranges from $5{\mu}s$ to $9{\mu}s$, and the
switching time ranges from $2{\mu}s$ to $4{\mu}s$.
\item The initial state is with empty stack, empty \verb|pc|, cleared 
\verb|mask| and cleared \verb|ir|.
\end{itemize}

@@

Choosing the \emph{maximal time sampling strategy}, the results of the
model checking show that both the schedulability and the correctness
properties hold in Scenarios (i-iii), but not in Scenario (iv). For
Scenario (iv), one counterexample of the schedulability is pictured in
Figure~\ref{f:counterexample}, where the third task (task$2$) misses
its deadline at the time $15ms$ that is marked by the tiny red
triangle. And the correctness fails in Scenario (iv) because of the
failure of schedulability.

@@

\subsection{Completeness of the Analysis}

>>>>

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathabx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumerate}
\usepackage{epic}
\usepackage{pstricks}
%\usepackage{llncsdoc}

@@

\subsection{Specification}
A Real-Time Maude module specifies a \emph{real-time rewrite theory}
${\cal R} = (\Sigma, E\cup A , IR, TR)$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of possibly conditional equations, and $A$ a
  set of equational axioms such as associativiy, commutativity and
  identity.  $(\Sigma, E\cup A)$ specifies the system's state space as
  an algebraic data type, and includes a built-in specification of a
  sort \verb|Time|.
\item $IR$ is a set of \emph{labeled conditional rewrite rules}
  specifying the system's local transitions, each of which has the
  form $[l]~:~t\rightarrow t'\mbox{ \textbf{if}
  }\bigwedge^n_{j=1}cond_j$, where each $cond_j$ is an equality
  $u_j=v_j$, and $l$ is a \emph{label}. Such a rule specifies an
  \emph{instaneous transition} from an instance of $t$ to the
  corresponding instance of $t'$, \emph{provided} the conditions hold.
\item $TR$ is a set of \emph{tick rules} $[l]~:~\{t\}\rightarrow\{t'\}
  \mbox{ \textbf{in time} }\tau\mbox{ \textbf{if} }cond$ that advance
  time in the \emph{entire} state $t$ by $\tau$ time units.
\end{itemize}

@@

\subsection{Scenarios}
\label{ss:results}
We use the following setting for our verification, which is from the 
statistics provided by our industrial partner:
\begin{itemize}
\item The interrupt cycle is $5ms$.
\item The scheduling time ranges from $5{\mu}s$ to $9{\mu}s$, and the
switching time ranges from $2{\mu}s$ to $4{\mu}s$.
\item The initial state is with empty stack, empty \verb|pc|, cleared 
\verb|mask| and cleared \verb|ir|.
\end{itemize}

@@

Instantiating our model on dense time domain and choosing the
\emph{maximal time sampling strategy}, the results of the model
checking show that both the schedulability and the correctness
properties hold in Scenarios (i-iii), but not in Scenario (iv). For
Scenario (iv), one counterexample of the schedulability is pictured in
Figure~\ref{f:counterexample}, where the third task (task$2$) misses
its deadline at the time $15ms$ that is marked by the tiny red
triangle. And the correctness fails in Scenario (iv) because of the
failure of schedulability.

@@

\subsection{Soundness and Completeness of the Analysis}
We call an analysis method \emph{sound} if any counterexample found
using such a method is a real counterexample of the question, and
\emph{complete} if the fact that no counterexample can be found using
such a method implies no counterexample exist for the question in
analysis. The soundness of our results is easy to check, simply by
examining the counterexamples found. But this is not the case for
completeness, since we choose instantiating our model on dense time
domain to make it more real but giving rise to an infinite state space
which is unfeasible to exhaust. In general, completeness of untimed
model checking cannot be achieved for any systems, any time sampling
stategies and any properties. However, \"Olveczky and Meseguer proved
the completeness of untimed temporal logic model checking, under the
maximal time sampling strategy, for a large class of real-time
systems:
\begin{theorem}
\label{t:completeness}
Given a \emph{time-robust} real-time rewrite theory $\cal R$, a set
$P$ of \emph{tick-stabilizing} atomic propositions, an LTL formula
$\Phi$ (excluding the \emph{next} operator $\bigcirc$) whose atomic
propositions are contained in $P$. The untimed temporal logic model
checking verifying $\Phi$ is \emph{complete} under the maximal time
sampling stategy.
\end{theorem}

To apply Theorem~\ref{t:completeness} on our analysis to show the
completeness, we only need to prove time-robustness of our model and
tick-stabilization of our defined atomic propositions,
\verb|taskTimeout| and \verb|scheduleCorrectly|. The latter is quite
trivial, while the former requires little adjustments to the model,
which collapse the possible finishing interval for some periodic task
into a single value that should be assigned to the possible maximum
computation time. Then we can apply Lemma~\ref{l:timerobustness} to
prove the time-robustness of our (adjusted) model.
\begin{lemma}
\label{l:timerobustness}
An object-oriented specification $\cal R$, with a conventional tick
rule, is \emph{time-robust} if the following conditions are satisfied
for all appropriate \emph{ground} terms $t$ and $r,r'$:

\noindent
(i) $mte(\delta(t,r))=mte(t)\dotdiv r$, for all $r\le mte(t)$;

\noindent
(ii) $\delta(t,0)=t$;

\noindent
(iii) $\delta(\delta(t,r),r')=\delta(t,r+r')$, for $r+r'\le mte(t)$;

\noindent
(iv) $mte(\sigma(l))=0$ for each ground instance $\sigma(l)$ of a
left-hand side of an instantaneous rewrite rule.
\end{lemma}

We achieve the same verification results for the adjusted model as for
the original one: schedulability and correctness hold in Scenarios
(i-iii) but fail in Scenario~(iv). The only thing remained to show the
completeness of our results is to prove the reasonability of the
adjustments, that is, if a set of tasks (including system ones) is
schedulable when the computation time of some task $\tau_i$ in it
equals to $C_i^{max}$, then when some jobs of $\tau_i$ become possible
to finish in time $r\le C_i^{max}$, the set of tasks remains still
schedulable. This is done by proving Lemma~\ref{l:max} and considering
\verb|scheduling| and \verb|switching| as having same and higher
priority than all periodic tasks.
\begin{lemma}
\label{l:max}
Assume the time $r_s$ at which a job of $\tau_i$ starts running is
fixed. The time $r_f$, at which this job finishes, reaches its maximum
if all job instances of tasks $\tau$'s, the priorities of which are
higher than or equal to that of $\tau_i$, running between $r_s$ and
$r_f$ finish in their maximum possible computation time.
\end{lemma}
